---
title: "Tukey HSD Tests for Cases per Neighborhood"
output:
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook is used to conduct if there is significant difference between COVID-19 cases and rate between neighborhood clusters as detected earlier in Python. 

```{r}
# Read in the data
data = read.csv("data/case_per_neighborhood.csv", head = TRUE, sep = ",")

data$cluster = factor(data$cluster)

# Show the first few rows of data
head(data, 3)
```

Boxplot of *case_per_capita* and *cluster*, with *case_per_capita* on the vertical axis:

```{r}
boxplot(case_per_capita~cluster,data=data,
        xlab="cluster", ylab="case_per_capita")
```

There seems to be noticeable difference in means of *case_per_capita* across the clusters. Let's see about *positive_rate*:

```{r}
boxplot(positive_rate~cluster,data=data,
        xlab="cluster", ylab="positive_rate")
```

First, one-way ANOVA tests will be conducted on *case_per_capita* to confirm if the means across clusters are statistically significant. The null hypothesis is that all means are equal.

```{r}
case_per_capita_aov = aov(case_per_capita~cluster,data=data)

summary(case_per_capita_aov)
```

At $\alpha$-level of 0.01,F-test p-value here is much smaller than that. Therefore, we reject the null hyphothesis of equal *case_per_capita* means between *cluster*. At least a pair of clusters has statistically significantly different *case_per_capita* means.

```{r}
TukeyHSD(case_per_capita_aov, conf.level = 0.99)
```

As Tukey HSD test confirms, there is statistically signficant difference between Cluster 1 and Cluster 0 and also between Cluster 1 and Cluster 2 while the test do not reject the null hypothesis for difference between Cluster 2 and Cluster 0. Cluster 1 has the lowest *case_per_capita* mean.

Now, even if Cluster 2 and Cluster 0 has probably equal means in *case_per_capita*. They might be different in health issues that led to the division into 2 clusters.

Boxplot of *obesity_percent* and *cluster*, with *obesity_percent* on the vertical axis:

```{r}
boxplot(obesity_percent~cluster,data=data,
        xlab="cluster", ylab="obesity_percent")
```

```{r}
obesity_aov = aov(obesity_percent~cluster,data=data)

summary(obesity_aov)
```

ANOVA test does confirm the difference in *obesity_percent* means across clusters.


```{r}
TukeyHSD(obesity_aov, conf.level = 0.99)
```

All clusters have statiscally different  *obesity_aov* with Cluster 1 still has the lowest while Cluster 2 has the highest mean.

Boxplot of *copd_rate* and *cluster*, with *copd_rate* on the vertical axis:

```{r}
boxplot(copd_rate~cluster,data=data,
        xlab="cluster", ylab="copd_rate")
```

```{r}
copd_aov = aov(copd_rate~cluster,data=data)

summary(copd_aov)
```

ANOVA test confirms the difference in *copd_rate* means across clusters.

```{r}
TukeyHSD(copd_aov, conf.level = 0.99)
```

While again, Cluster 1 demonstrates a significant lower *copd_rate* mean between other clusters, Cluster 0 and Cluster 2 does not have any statistically difference in this health aspect.

Boxplot of *asthma_percent* and *cluster*, with *asthma_percent* on the vertical axis:

```{r}
boxplot(asthma_percent~cluster,data=data,
        xlab="cluster", ylab="asthma_percent")
```

```{r}
asthma_aov = aov(asthma_percent~cluster,data=data)

summary(asthma_aov)
```

ANOVA test confirms the difference in *asthma_percent* means across clusters.

```{r}
TukeyHSD(asthma_aov, conf.level = 0.99)
```

Interestingly, only Cluster 2 exhibits a statiscally higher *asthma_percent* mean than other clusters. Cluster 1 and Cluster 0 does not have a significant difference between their *asthma_percent*.

Boxplot of *heart_attack_percent* and *cluster*, with *heart_attack_percent* on the vertical axis:

```{r}
boxplot(heart_attack_percent~cluster,data=data,
        xlab="cluster", ylab="heart_attack_percent")
```

```{r}
heart_aov = aov(heart_attack_percent~cluster,data=data)

summary(heart_aov)
```

ANOVA test confirms the difference in *heart_attack_percent* means across clusters.

```{r}
TukeyHSD(heart_aov, conf.level = 0.99)
```

We see something similar again in  *heart_attack_percent*, Cluster 1 again has signficantly lower  *heart_attack_percent* mean than the other 2 while between Cluster 0 and Cluster 2 they do not have a statistically difference.

In conclusion, the lower *case_per_capita* in Cluster 1 could be driven by the fact that it has lower *obesity_percent*, *copd_rate*, and *heart_attack_percent* than other clusters.

Though Cluster 0 and Cluster 2 do not have a statistically difference in *case_per_capita* means, they are different because of their difference in *obesity_percent* and *asthma_percent* .








